Automated Dynamic Slice Backsolving

### 2014-03-21 docs/paper/fsuthesis/thesis-template/paper.tex
safety commit
looking at todos, working on contributions like design and implementation

\begin{acknowledgments}
SFS
% Dr. Wang
% Bret Whissel
\end{acknowledgments}

@@

@@

\section{Contributions}

@@

TODO add loop invariants experiments

@@

SAGE uses symbolic execution, and prunes paths taken with a propietary "generational search" algorithm. This algorithm prioritizes generating new test cases (TODO does it draw on the test case reduction and expansion algorithms discussed in that one paper???). It takes a set of all constraints and tries negating each constraint independently to generate a new test case. SAGE then symbolically re-executes the constraint set using TruScan [TODO CITE] [Q]. Similarly, s2e uses QEMU for instruction set emulation, and then appears to be a strong trend in white-box fuzzing. TODO ADV/DIS?

@@

%\section{Grey-box Evolutionary Fuzzing}

%TODO

@@

%\section{TODO Difficulties and Limitations}

%TODO Path explosion. Code coverage. Complex file formats which mean you create shallow errors over and over again.

@@

\section{Intel PIN}

@@

@@

Stubble architecture, describe
why intel pin and not dynamo rio, valgrind, etc.

@@

@@

Pintool, python specifics. Binarys, RE

\section{Limitations}

@@

TODO explain how brute force fuzzers or regular oracles would solve, add their expected number of attempts to graph.
(TODO time smaller oracle attacks)
TODO attacks which transform input first to show ability to backsolve

@@

%We have implemented a proof of concept fuzzer which uses dynamic taint analysis to record the operations performed on user input, then reverses these operations to solve for new paths to explore. The fuzzer currently records implicit and explicit taint flows between memory and registers. When the EFLAGS register is modified, it is checked whether a comparison instruction with tainted operands was the modifier, and if so this is also recorded. After an execution it publishes input, output, taint, and mutation data to a driver program, which reasons about possible user input locations and values one byte at a time. This means when branch decisions are determined by comparisons at the byte level, the fuzzer can guarantee a new path is taken, provided it can accurately backsolve information about the transformations performed on the byte.

@@

%We assume ASLR is disabled for the binary to be fuzzed. This allows the program to easily track and compare execution paths. We have implemented the technique for the x86-32 architecture, because x86 is widespread, and vulnerabilities resulting from unsafe C code should persist across binaries, regardless of both the size and flavor of architecture (32 vs. 64-bit, ARM vs. x86 vs. PowerPC etc.), since compilers try to adhere to the C standard and vulnerabilities like buffer overflows exist at a C source code level. We also assume the presence of an algorithm which will detect an exploitable condition, provided it is given the proper binary and input to the binary, and an initial test case. The program concretely executes the binary once with the initial test case, recording taint propagation during execution. Stdin, stdout, and stderr are recorded for each run, for later analysis. After the initial run, the program automatically starts solving each byte by backsolving to try and find new paths.

%TODO in above make sure can catch stderr segfaults, segfaults go to shell's stderr, also a signal

@@

\section{Testing Harness}

Fuzzers require intuitive but powerful test harnesses, so that users can easily control how a binary is fuzzed and understand the results of fuzz testing. Currently, stubble records input, output, taint for each test executed, appending an auto-incrementing id to log files and separating results at the directory level. Stubble also keeps track of the mutations which have already been explored. A driver program keeps track of the observed taint which has already been explored, and selects a new byte to mutate. Users can grep for interesting conditions like segfaults or unexpected output, then cross reference the test number with the observed taint flow, input file, particular mutation tested.

@@

>>>>

\begin{acknowledgments}
SFS
% Dr. Wang
% Bret Whissel
% Salwan
\end{acknowledgments}

@@

%TODO add RE protocol to abstract and intro once implemented

@@

\section{Contribution}

@@

@@

SAGE uses symbolic execution, and prunes paths taken with a propietary "generational search" algorithm. This algorithm prioritizes generating new test cases. It takes a set of all constraints and tries negating each constraint independently to generate a new test case. SAGE then symbolically re-executes the constraint set using TruScan [TODO CITE] [Q]. Similarly, s2e uses QEMU for instruction set emulation, and then appears to be a strong trend in white-box fuzzing. TODO ADV/DIS?

@@

%\section{Grey-box Evolutionary Fuzzing} todo

@@

@@

\section{Intel Pin}

% TODO check if mention Pin compiles from x86 to x86 adding pintool

@@

%TODO this tech helps with complex file formats which mean you create shallow errors over and over again.

@@

TODO Stubble architecture, describe
why intel pin and not dynamo rio, valgrind, etc.

Stubble consists of a pintool which implements DTA, and a Python driver which keeps track of test cases, input and output recorded during execution, and employs dynamic slice backsolving to drive mutation using taint analyses provided by the DTA pintool. Pintools are arbitrary C++ which Pin compiles and inserts into a given binary during instrumentation. The driver accepts a valid x86 binary and a seed test case. From the seed test case, the driver runs the pintool on the binary with the seed test case as input. While the binary runs, the pintool notes when taint in introduced or removed from memory or from registers. If a memory location or a register is tainted, then any operations which move this data elsewhere or use this data in an operation spread taint to the appropriate areas. Moving untainted data into a previously tainted location removes taint from the system. For instance, if the eax register is tainted, then adding eax and ebx and storing the result in ebx will taint the ebx register. However, should an operation move the value of ebx to eax, then eax will no longer be tainted.

In addition, whenever stubble encounters an instruction that modifies the EFLAGS register, such as the cmp instruction, it is checked to see whether either of the operands is tainted. When this is the case, user input has some level of control over the EFLAGS register, so the appropriate flag in the register is marked tainted. When a conditional jump instruction is reached, the corresponding flag of which is tainted, stubble begins to backsolve with the goal of providing a test case to explore the other side of the branch. After an execution it publishes input, output, taint, and mutation data to the driver, which reasons about possible user input locations and values one byte at a time. This means when branch decisions are determined by comparisons at the byte level, the fuzzer can guarantee a new path is taken, provided it can accurately backsolve information about the transformations performed on the byte.

%The fuzzer currently records implicit and explicit taint flows between memory and registers. When the EFLAGS register is modified, it is checked whether a comparison instruction with tainted operands was the modifier, and if so this is also recorded.

%We also assume the presence of an algorithm which will detect an exploitable condition, provided it is given the proper binary and input to the binary, and an initial test case. The program concretely executes the binary once with the initial test case, recording taint propagation during execution. Stdin, stdout, and stderr are recorded for each run, for later analysis. After the initial run, the program automatically starts solving each byte by backsolving to try and find new paths.

@@

We assume ASLR is disabled for the binary to be fuzzed. This allows the program to easily track and compare execution paths. We have implemented the technique for the x86-32 architecture, because Intel Pin works on a select few architectures, of which x86 is the most prevalent. Since x86 is quite widespread, and vulnerabilities resulting from unsafe C code should persist across binaries, regardless of both the size and flavor of architecture (32 vs. 64-bit, ARM vs. x86 vs. PowerPC etc.), this should be no issue. Compilers try to adhere to the C standard and vulnerabilities like buffer overflows exist at a C source code level.

Fuzzers require intuitive but powerful test harnesses, so that users can easily control how a binary is fuzzed and understand the results of fuzz testing. Currently, stubble records input, output, taint for each test executed, appending an auto-incrementing id to log files and separating results at the directory level. Stubble also keeps track of the mutations which have already been explored. Users can grep for interesting conditions like segfaults or unexpected output, then cross reference the test number with the observed taint flow, input file, and particular mutation tested.

@@

Pintool, python specifics. Binarys, RE
how i implemented DTA

\section{Challenges}

% undertainting examples, general DTA issues and how we solved.

@@

(TODO time smaller oracle attacks)
TODO attacks which transform input first to show ability to backsolve

@@

@@

%TODO in above make sure can catch stderr segfaults, segfaults go to shell's stderr, also a signal, test this

@@

@@

\chapter{Conclusions}

\section{Applications}

TODO ways this technique could be used.
TODO talk about why parallelization is good.

\section{Future Work}
